Fixing Vulnerabilities Potentially Hinders Maintainability

### 2020-05-28 paper/icsme20.tex
minor fixes
pass in implications section

\section{Study Implications}\label{sec:implications}

This section presents the implications that we retrieved from this
study and think are important to consider in the future by developers.
Our results exhibit evidence that developers may have to
reduce maintainability for the sake of security. We understand that developers
should be able to patch and produce secure code without hindering the maintainability of
their projects. But there are still some concerns that need to be addressed
and that this study tries to bring awareness for:
%
\\\textit{\textbf{Tools for Patch Risk Detection:}}
	Design debt of one guideline can lead
	to severe impacts on the software quality~\cite{10.1145/1985362.1985366}. 
	Thus, tools like Better Code Hub and static analysis (e.g., SonarQube, Codacy,
	ESLint, Infer, and more) should be leveraged to 
	assess the risk of patching vulnerabilities---alongside pull
	requests/code reviews. While static code analysis
	may be daunting due to the number of rules and different effects
	on maintainability, BCH uses the top-$10$ of the guidelines
	with the highest effect on maintainability. Yet, static analysis 
	can be used to complement the source code analysis, introduce
	the capability of vulnerability detection, and, predict and
	reduce the risk of hindering maintainability.
%
\\\textit{\textbf{Computer Curricula Needs to be Updated:}}
	Software Developers lack the knowledge to create secure and
	maintainable code. Mainly because academic curricula is not yet prepared
	to properly educate them. Universities should do their part on making
	this possible. Online services such as BCH have a great potential to help
	developers to be more aware of the maintainability issues introduced by their
	changes. Thus, developers can 
	put more work into improving their code maintainability and avoiding the risk
	of introducing new vulnerabilities while patching older ones.
%	
\\\textit{\textbf{Secure Programming Languages Wanted:}} Programming Languages 
	should provide new design patterns to easily implement security weaknesses 
	without endangering maintainability.
	Ultimately, new programming languages, both secure and maintainable by design,
	may be designed to help developers be highly productive when writing and
	maintaining secure applications. One example is the need for designing new
	authorization mechanisms since these are one of the most complex security
	features to implement.
%
\\\textit{\textbf{Poor Coding Practices Documentation:}}
	Software documentation must provided to developers with the
	best practices to implement secure and maintainable software. Unmaintainable
	code is difficult to test, analyze and re-use. Software companies must
	do an effort to provide these artifacts to their developers, in order,
	to avoid the introduction of new vulnerabilities while patching older ones.

\section{Threats to Validity}\label{sec:threats}
%
This section presents the potential threats to the validity.
%
\\\textit{\textbf{Construct:}} The formula was inferred based 
on the BCH's reports. The high amount of different projects and backgrounds 
may require other maintainability standards. However, BCH does use a 
representative benchmark of closed and open-source software projects to compute 
the thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012}.
Maintainability is computed as the mean of all guidelines. Different software versions (vulnerable/fixed) of one vulnerability may have the same overal score
and still be affected by different guidelines.
%
\\\textit{\textbf{Internal:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed by them
to fix security flaws. This approach discards security patches that were
not explicit in commits messages. Pontas et. al\cite{10.1109/MSR.2019.00064} dataset
has patches that were manually mined in software repositories and, consequently,
wrong associations may have been performed between vulnerabilities and code snippets. 
The dataset also included floss-refactorings. Both cases were manually inspected
and deleted from the dataset. Thus, this study does not consider that a
developer may perform a patch quickly and later proceed to the refactor.
Baseline commits are retrieved randomly by selecting one commit with an 
approximate size and from the same project of a given security patch. 
This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style and software contribution policies which is not evaluated in this study.
Furthermore, this evaluation considers that $969$ regular commits---any kind
of commit---are enough to
alleviate random irregularities in the maintainability differences of the
baseline. 
\\\textit{\textbf{External:}} We use a dataset that only comprises patches of open-source software.
However, the methodology requires to access data that is not publicly available.
Thus, our findings may not extend to non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. Our approach does not consider 
security patches in any other languages but English.

>>>>

\section{Study Implications}\label{sec:implications}
%
Our results show evidence that developers may have to reduce maintainability for the 
sake of security. We argue that developers should be able to patch and produce secure 
code without hindering the maintainability of their projects. But there are still  
concerns that need to be addressed and that this study brings awareness for:
%
\\\textit{\textbf{Tools for Patch Risk Detection:}}
	Design debt of one guideline can lead
	to severe impacts on the software quality~\cite{10.1145/1985362.1985366}. 
	Thus, tools like Better Code Hub and static analysis (e.g., SonarQube, Codacy,
	ESLint, Infer, and more) should be leveraged to 
	assess the risk of patching vulnerabilities---alongside pull
  requests/code reviews. \Rui{what? (till the ned)-->} While static code analysis
	may be daunting due to the number of rules and different effects
	on maintainability, BCH uses the top-$10$ of the guidelines
	with the highest effect on maintainability. Yet, static analysis 
	can be used to complement the source code analysis, introduce
	the capability of vulnerability detection, and, predict and
  reduce the risk of hindering maintainability.
  \Rui{static analysis vs. source coode?!}
%
\\\textit{\textbf{Computer Curricula Needs to be Updated:}}
	We argue that software Developers lack the knowledge to create both secure and
	maintainable code. Mainly because academic curricula is not yet prepared
	to properly educate them. Universities should do their part on making
	this possible. \Rui{connection to curricula is not clear-->} Online services such as BCH have a great potential to help
	developers to be more aware of the maintainability issues introduced by their
	changes. Thus, developers can 
	put more work into improving their code maintainability and avoiding the risk
	of introducing new vulnerabilities while patching older ones.
%	
\\\textit{\textbf{Secure Programming Languages Wanted:}} Programming Languages 
	should provide new design patterns to easily implement security weaknesses 
	without endangering maintainability.
	Ultimately, new programming languages, both secure and maintainable by design,
	may be designed to help developers be highly productive when writing and
	maintaining secure applications. One example is the need for designing new
	authorization mechanisms since these are one of the most complex security
	features to implement.
%
\\\textit{\textbf{Poor Coding Practices Documentation:}}
  \Rui{too generic?}
	Software documentation must be provided to developers with the
	best practices to implement secure and maintainable software. Unmaintainable
	code is difficult to test, analyze and re-use. Software companies must
	do an effort to provide these artifacts to their developers, in order,
	to avoid the introduction of new vulnerabilities while patching older ones.

\section{Threats to Validity}\label{sec:threats}
%
This section presents the potential threats to the validity.
%
\\\textit{\textbf{Construct:}} The formula was inferred based 
on the BCH's reports. The high amount of different projects and backgrounds 
may require other maintainability standards. However, BCH does use a 
representative benchmark of closed and open-source software projects to compute 
the thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012}.
Maintainability is computed as the mean of all guidelines. Different software versions (vulnerable/fixed) of one vulnerability may have the same overal score
and still be affected by different guidelines.
%
\\\textit{\textbf{Internal:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed by them
to fix security flaws. This approach discards security patches that were
not explicit in commits messages. Pontas et. al\cite{10.1109/MSR.2019.00064} dataset
has patches that were manually mined in software repositories and, consequently,
wrong associations may have been performed between vulnerabilities and code snippets. 
The dataset also included floss-refactorings. Both cases were manually inspected
and deleted from the dataset. Thus, this study does not consider that a
developer may perform a patch quickly and later proceed to the refactor.
Baseline commits are retrieved randomly by selecting one commit with an 
approximate size and from the same project of a given security patch. 
This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style and software contribution policies which is not evaluated in this study.
Furthermore, this evaluation considers that $969$ regular commits---any kind
of commit---are enough to
alleviate random irregularities in the maintainability differences of the
baseline. 
%
\\\textit{\textbf{External:}} We use a dataset that only comprises patches of open-source software.
However, the methodology requires to access data that is not publicly available.
Thus, our findings may not extend to non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. Our approach does not consider patches in any other languages but English.
